---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit THIS file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  results = "asis",
  echo = TRUE,
  comment = "#>",
  fig.path = "vignettes/figures/README-",
  out.width = "100%"
)
IS_README <- TRUE
library(datelife)
```
<!-- badges: start -->

<!-- [![R build status](https://github.com/phylotastic/datelife/workflows/R-CMD-check/badge.svg)](https://github.com/phylotastic/datelife/actions) -->
![GitHub R package version](https://img.shields.io/github/r-package/v/phylotastic/datelife?color=green)
[![Build Status](https://travis-ci.org/phylotastic/datelife.svg)](https://travis-ci.org/phylotastic/datelife)
[![codecov](https://codecov.io/gh/phylotastic/datelife/branch/master/graph/badge.svg)](https://codecov.io/gh/phylotastic/datelife)
[![Github Open Issues](https://img.shields.io/github/issues-raw/phylotastic/datelife.svg)](https://github.com/phylotastic/datelife/issues)
[![Github Closed Issues](https://img.shields.io/github/issues-closed-raw/phylotastic/datelife.svg)](https://github.com/phylotastic/datelife/issues?q=is%3Aissue+is%3Aclosed)
[![DOI](https://zenodo.org/badge/23036/phylotastic/datelife.svg)](https://zenodo.org/badge/latestdoi/23036/phylotastic/datelife)
[![NSF-1458603](https://img.shields.io/badge/NSF-1458603-white.svg)](https://nsf.gov/awardsearch/showAward?AWD_ID=1458603)
[![NSF-0905606](https://img.shields.io/badge/NSF-0905606-white.svg)](https://nsf.gov/awardsearch/showAward?AWD_ID=0905606)
[![NSF-1458572](https://img.shields.io/badge/NSF-1458572-white.svg)](https://nsf.gov/awardsearch/showAward?AWD_ID=1458572)

<!-- badges: end -->

# Welcome to DateLife's GitHub repository!

Get a phylogenetic tree with branch lengths proportional to geologic time (aka a
_**chronogram**_) of any two or more lineages of interest to you: use the `datelife`
R package or go to [www.datelife.org](http://datelife.org/query/) to make
a query of chronograms available for your lineages in the [Open Tree of Life](https://tree.opentreeoflife.org/curator)'s
tree store.

You can also date a phylogenetic tree of your own making (or choosing one from the literature), using node ages from chronograms queried with `datelife`
and the congruification method ([Eastman et al. 2013](http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12051/abstract))
implemented with _geiger_ ([Harmon et al. 2008](http://bioinformatics.oxfordjournals.org/content/24/1/129.short)).


`datelife` has been developed as part of the [phylotastic](http://phylotastic.org/) (NSF-funded) project, and is still under rapid development.

- [Installation](#installation)
- [Quick intro](#quick-intro)
- [Citation](#citation)
- [Feedback and info for developers](#feedback)
- [License](#license)


## Installation

Get the stable version from [CRAN](https://cran.r-project.org/web/packages/datelife/index.html):

```{r eval = FALSE}
install.packages("datelife")
```

Or, install the development version from the GitHub repository:

```{r eval = FALSE}
devtools::install_github("phylotastic/datelife")
```

## Quick intro

You can update the cached objects with the `update_all_cached()` function.


## Citation

If you use `datelife` for a publication, please cite the R package and the accompanying paper:

```{r eval = TRUE, echo = FALSE}
print(citation("datelife"), bibtex = FALSE, style = "html")
```

You can get these citations and the bibtex entry with:
```{r eval = FALSE}
citation("datelife")
toBibtex(citation("datelife"))
```

## Feedback and Information for Developers

All comments, ideas and questions about `datelife` are encouraged. You are welcome
to post an issue [here](https://github.com/phylotastic/datelife/issues/new), or
to make a [pull request](https://github.com/phylotastic/datelife/pulls) if you
want to contribute with code directly.

Package and function documentation was generated with [roxygen2](https://CRAN.R-project.org/package=roxygen2):

```{r eval = FALSE}
roxygen2::roxygenise()
```

We used the package [lintr]() to check for coding style:

```{r eval = FALSE}
lintr::lint_package()
```

Code coverage was tested with the package [covr](https://CRAN.R-project.org/package=covr):

```{r eval = FALSE}
cov <- covr::package_coverage()

usethis::use_data(cov, overwrite = TRUE)
```

You can see an interactive report of testing coverage:

```{r eval = FALSE}
covr::report(cov)
```

And, find code with zero coverage:

```{r eval = FALSE}
covr::zero_coverage(cov)
```

To create the CRAN release, first run a `R CMD check` from your terminal.
For that, change directories to the one above your package:

```{bash, eval = FALSE}
cd ../
```

Generate a tar ball for your package by running `R CMD build`:

```{bash, eval = FALSE}
R CMD build datelife
```

Finally, run `R CMD check` on the tar ball that you just generated:

```{bash, eval = FALSE}
R CMD check datelife_0.5.0.tar.gz
```


### Generating `datelife`s hexsticker

The chronogram was generated with:

```{r eval = FALSE}
phylo_sdm <- datelife::datelife_search(input = c("Delphinus_delphis", "Gallus gallus", "elephas Maximus", "felis_catus", "homo-sapiens"),
                                       summary_format = "phylo_sdm")
phylo_sdm$tip.label_original <- phylo_sdm$tip.label
phylo_sdm$tip.label <- c("elephant", "woman_cartwheeling", "dolphin", "cat", "rooster")
class(phylo_sdm) <- "phylo"
```

And plotted with:

```{r eval = FALSE}
BiocManager::install("ggtree")
install.packages("emojifont")
library(emojifont)
library(ggimage)
p <- ggtree::ggtree(phylo_sdm, size=0.5, color='black') +  
  ggtree::geom_tiplab(parse="emoji", size=8, vjust=0.01) +
  ggtree::theme_tree(bgcolor = "#00640050")
svglite::svglite("inst/figures/phylo-sdm-emoji.svg", width = 10, height = 7)
print(p)
dev.off()
```


```{r eval = FALSE}
install.packages("hexSticker")

s <- hexSticker::sticker(~print(p),
          package="datelife", p_size=20, s_x=.8, s_y=1, s_width=1.4, s_height=1.2,
          filename="inst/figures/datelife-hex.png")
          
```

### Rendering the vignettes

```{r eval = FALSE}
knitr::knit("vignettes/Getting_started_with_datelife.Rmd")
```

## License

This package is free and open source software, licensed under GPL.
