---
title: "Tools for package development"
output:
  rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Tools for package development}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(datelife)
```

# Tools for package development
And other tips that I used to get `datelife` on CRAN

https://www.r-bloggers.com/2020/12/top-10-tips-to-make-your-r-package-even-more-awesome/

## Dependencies on packages outside of CRAN

### Bioconductor packages:

I was having trouble declaring Bioconductor packages.

Most people suggested that adding "BiocViews:" to the Description file would solve
the issue. But it did not, and I kept on finding unanswered posts about packages
being rejected fromx CRAN because of dependencies from Bioconductor using the "biocviews"
trick.
So I did a reverse dependencie search on Bioconductor packages. Surely at least
one package in CRAN must have a dependency from Bioconductor. Indeed, the
[msa](https://bioconductor.org/packages/release/bioc/html/msa.html) package has
[bio3d](https://CRAN.R-project.org/package=bio3d) as a reverse dependency:

```r
devtools::revdep("msa")

#> [1] "bio3d"
```

Going into the [CRAN mirror of the bio3d package on GitHib](https://github.com/cran/bio3d),
I found their solution:

```r
if (!requireNamespace("msa", quietly=TRUE)) {
  stop("msa package missing: Please install it from Bioconductor, see: ?BiocManager::install")
}
```


Calling Bioconductor packages conditionally allows us to move them into "Suggests"
on the DESCRIPTION file. I found out you also have to add them to "Remotes":

```
Suggests:
    msa,
    Biostrings
Remotes:
    bioc::release/msa,
    bioc::release/Biostrings
```

The Remotes part had to be removed for CRAN submission. Because some of my tests
depend on these packages being installed, I added  helper file with a call to
package installation of msa and Biostrings if they are not already installed.




Note `checking CRAN incoming feasibility   Maintainer:`

It is safe to ignore this note:
https://stackoverflow.com/questions/23829978/checking-cran-incoming-feasibility-note-maintainer

Automated R CMD checks with GitHub actions:

```
remotes::install_github("aaronpeikert/repro")

repro::check_git()
repro::check_github_ssh()
usethis::create_github_token()

gitcreds_set() Choose option 2) Replace credentials. It will prompt your for the
new token, and you will be all set.

usethis::use_github_actions()
```
