---
title: "Case study: the Fringiliidae"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case study: the Fringiliidae}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(datelife)
```

## 1. Processing Fringiliidae names

Get all species names belonging to the family with the flag `get_spp_from_taxon = TRUE`:

```r
dq <- datelife::make_datelife_query(input = "fringilidae",               
                                   get_spp_from_taxon = TRUE)

#> ... Making a DateLife query.
#> ... Phylo-processing 'input'.
#> 'input' is not a phylogeny.
#>   |=========================================================| 100%
#>   |=========================================================| 100%
#>   |=========================================================| 100%
#>   |=========================================================| 100%
#>   |=========================================================| 100%
#> Working with the following taxa: 
#>	Crithagra sulphurata | Crithagra citrinelloides | Crithagra mozambica | Crithagra capistrata | Crithagra burtoni | Crithagra rufobrunnea | Crithagra flaviventris | Callacanthis burtoni | Chaunoproctus ferreorostris | Rhynchostruthus socotranus | Chlorophonia cyanea | Chlorophonia occipitalis | Chlorophonia flavirostris | Akialoa obscura | Euphonia gouldi | Euphonia plumbea | Euphonia elegantissima | Euphonia luteicapilla | Euphonia chrysopasta | Euphonia fulvicrissa | Euphonia hirundinacea | Euphonia chlorotica | Euphonia cyanocephala | Euphonia rufiventris | Euphonia minuta | Euphonia cayennensis | Euphonia xanthogaster | Euphonia violacea | Euphonia pectoralis | Euphonia musica | Euphonia anneae | Euphonia affinis | Euphonia laniirostris | Euphonia finschi | Psittirostra psittacea | Urocynchramus pylzowi | Telespiza cantans | Telespiza ultima | Palmeria dolei | Oreomystis bairdi | Loxops coccineus | Loxops mana | Loxops caeruleirostris | Himatione sanguinea | Rhodopechys sanguineus | Paroreomyza montana | Pseudonestor xanthophrys | Vestiaria coccinea | Bucanetes githagineus | Hemignathus parvus | Hemignathus wilsoni | Hemignathus virens | Hemignathus munroi | Hemignathus flavus | Hemignathus kauaiensis | Hemignathus stejnegeri | Loxioides bailleui | Eremopsaltria mongolica | Rhodospiza obsoleta | Peucedramus taeniatus | Cyanerpes caeruleus | Cyanerpes cyaneus | Cyanerpes nitidus | Cyanerpes lucidus | Fringillaria goslingi | Fringillaria striolata | Fringillaria impetuani | Fringillaria poliopleura | Fringillaria tahapisi | Hemispingus atropileus | Hemispingus verticalis | Hemispingus auricularis | Hemispingus calophrys | Hemispingus rufosuperciliaris | Hemispingus superciliaris | Hemispingus trifasciatus | Hemispingus piurae | Hemispingus xanthophthalmus | Hemispingus parodii | Hemispingus goeringi | Hemispingus melanotis | Hemispingus frontalis | Hemispingus reyi | Rowettia goughensis | Embernagra platensis | Schoeniclus rusticus | Schoeniclus aureolus | Schoeniclus elegans | Schoeniclus siemsseni | Schoeniclus spodocephala | Schoeniclus yessoensis | Tersina viridis | Orthogonys chloricterus | Buarremon apertus | Buarremon assimilis | Buarremon atricapillus | Buarremon basilicus | Buarremon virenticeps | Buarremon torquatus | Peucaea aestivalis | Peucaea cassinii | Peucaea botterii | Phaeothlypis fulvicauda | Donacospiza albifrons | Platyspiza crassirostris | Melopyrrha nigra | Carduelis citrinella | Carduelis corsicana | Carduelis carduelis | Uragus sibiricus | Melamprosops phaeosoma | Hesperiphona vespertina | Pyrrhoplectes epauletta | Haemorhous mexicanus | Haemorhous purpureus | Haemorhous cassinii | Kozlowia roborowskii | Serinus syriacus | Serinus albogularis | Serinus atrogularis | Serinus citrinipectus | Serinus citrinella | Serinus pusillus | Serinus canaria | Serinus alario | Serinus striolatus | Serinus melanochrous | Serinus whytii | Serinus mennelli | Serinus reichardi | Serinus scotops | Serinus flavivertex | Serinus totta | Serinus estherae | Serinus hypostictus | Serinus leucopygius | Serinus thibetanus | Serinus gularis | Serinus dorsostriatus | Serinus canicollis | Serinus serinus | Linaria flavirostris | Linaria cannabina | Chloris monguilloti | Chloris sinica | Chloris spinoides | Chloris ambigua | Chloris chloris | Spinus yarrellii | Spinus spinus | Spinus olivaceus | Spinus uropygialis | Spinus siemiradzkii | Spinus atriceps | Spinus dominicensis | Spinus xanthogastrus | Spinus magellanicus | Spinus pinus | Spinus notatus | Spinus tristis | Spinus spinescens | Spinus crassirostris | Spinus barbatus | Spinus atratus | Spinus lawrencei | Spinus psaltria | Spinus cucullatus | Neospiza concolor | Loxia curvirostra | Loxia pytyopsittacus | Loxia leucoptera | Loxia scotica | Eophona migratoria | Eophona personata | Haematospiza sipahi | Mycerobas icterioides | Mycerobas melanozanthos | Mycerobas affinis | Mycerobas carnipes | Pinicola subhimachala | Pinicola enucleator | Linurgus olivaceus | Leucosticte sillemi | Leucosticte arctoa | Leucosticte australis | Leucosticte nemoricola | Leucosticte tephrocotis | Leucosticte atrata | Leucosticte brandti | Carpodacus vinaceus | Carpodacus rodopeplus | Carpodacus davidianus | Carpodacus rhodochlamys | Carpodacus rubescens | Carpodacus synoicus | Carpodacus severtzovi | Carpodacus rodochroa | Carpodacus erythrinus | Carpodacus roseus | Carpodacus nipalensis | Carpodacus trifasciatus | Carpodacus rubicilloides | Carpodacus pulcherrimus | Carpodacus thura | Carpodacus eos | Carpodacus rubicilla | Carpodacus grandis | Carpodacus edwardsii | Carpodacus puniceus | Carpodacus formosanus | Coccothraustes coccothraustes | Acanthis hornemanni | Acanthis cabaret | Acanthis flammea | Pyrrhula nipalensis | Pyrrhula erythaca | Pyrrhula erythrocephala | Pyrrhula murina | Pyrrhula leucogenis | Pyrrhula aurantiaca | Pyrrhula pyrrhula | Fringilla montifringilla | Fringilla teydea | Fringilla coelebs | Fringilla polatzeki | Calcarius lapponicus | Calcarius pictus | Calcarius ornatus | Calcarius mccownii | Emberiza variabilis | Emberiza hortulana | Emberiza pusilla | Emberiza cioides | Emberiza tristrami | Emberiza cabanisi | Emberiza jankowskii | Emberiza cineracea | Emberiza rutila | Emberiza sahari | Emberiza buchanani | Emberiza chrysophrys | Emberiza capensis | Emberiza cia | Emberiza schoeniclus | Emberiza stewarti | Emberiza godlewskii | Emberiza pallasi | Emberiza leucocephalos | Emberiza socotrana | Emberiza melanocephala | Emberiza bruniceps | Emberiza koslowi | Emberiza cirlus | Emberiza flaviventris | Emberiza affinis | Emberiza fucata | Emberiza citrinella | Emberiza sulphurata | Emberiza caesia | Plectrophenax nivalis | Plectrophenax hyperboreus | Melophus lathami | Cyanoloxia glaucocaerulea | Cyanoloxia cyanoides | Cyanoloxia rothschildii | Caryothraustes canadensis | Caryothraustes poliogaster | Cyanocompsa brissonii | Cyanocompsa parellina | Passerina caerulea | Passerina cyanea | Passerina amoena | Passerina ciris | Passerina leclancherii | Passerina rositae | Passerina versicolor | Parkerthraustes humeralis | Spiza americana | Pheucticus aureoventris | Pheucticus ludovicianus | Pheucticus melanocephalus | Pheucticus chrysogaster | Pheucticus tibialis | Cardinalis cardinalis | Cardinalis sinuatus | Cardinalis phoeniceus | Porphyrospiza caerulescens | Rhodothraupis celaeno | Periporphyrus erythromelas
#> DateLife query done!
```

Inspect the structure of the output:

```r
str(dq)

#> List of 3
#>  $ cleaned_names: Named chr [1:289] "Crithagra sulphurata" "Crithagra #> citrinelloides" "Crithagra mozambica" "Crithagra capistrata" ...
#>   ..- attr(*, "names")= chr [1:289] "Fringillidae1" "Fringillidae2" #> "Fringillidae3" "Fringillidae4" ...
#>  $ ott_ids      : Named int [1:289] 903281 328908 1083721 901359 318589 #> 455546 903271 405220 242220 455541 ...
#>   ..- attr(*, "names")= chr [1:289] "Crithagra sulphurata" "Crithagra #> citrinelloides" "Crithagra mozambica" "Crithagra capistrata" ...
#>  $ phy          : logi NA
#>  - attr(*, "class")= chr "datelifeQuery"
```

We have 289 species names in the family Fringiliidae and we provided no input tree so `$phy` is `NA`.

## 2. Search the species names across a chronogram database

Load the prebuilt chronogram database:

```r
utils::data("opentree_chronograms", package = "datelife")
```

```r
dr <- datelife::get_datelife_result(input = dq, cache = opentree_chronograms)

#> ... Searching the chronogram database.
#> Search done!
#>
#> Input taxon names were found in 13 chronograms.
```

Results from the chronogram search are returned as a `datelifeResult` object:

class(dr)

