---
title: "Getting started"
output:
  rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(datelife)
```

## What can you do with `datelife`?

The aim of the `datelife` R package is to allow researchers and the general audience to obtain open scientific data on ages for their organisms of interest. These age data is extracted from dated phylogenetic trees (chronograms) that have been published and peer-reviewed in association with a scientific article in an indexed journal. As such, the age data provided by `datelife` constitutes state-of-the-art, peer-reviewed, public scientific knowledge, that can be accessed by experts and non-experts in the field alike.

As a `datelife` user, you can get age data for your organisms of interest in several forms:

- **MRCA**: Get the age of the most recent common ancestor (MRCA) of your organisms.

- **Original chronogram**: Get the chronograms from the original publications, pruned to contain your organisms only.

- **Biggest chronogram**: Get the original chronogram with the most organisms of interest present as tips.

- **Summary chronogram**: Obtain a single chronogram, summarized using data from the original chronograms. The summary chronogram can be calculated using different methods.

You can also apply the obtained age data to date a phylogenetic tree topology of your organisms of interest.

## 1. Processing your organism names

The function `make_datelife_query` processes the taxon names of your organisms of interest and structures them as a `datelifeQuery` object.

A `datelifeQuery` object is a list of three elements:

- `$cleaned_names`: a character vector of "cleaned" taxon names, 
- `$ott_ids`: a numeric vector of OpenTree Taxonomic identifiers (OTT ids), and 
- `$phy`: an input tree as a `phylo` object or `NA` if none was provided.

This allows us to straightforwardly use the processed taxon names later to perform a search for ages across the `datelife` chronogram database as well as other downstream analyses. 

There are two alternative ways of processing organism names with `datelife`:

### a) From a set of organism taxonomic names
You can provide the taxon names as a character vector or as a single comma separated string of names.
For example, running the following:

```{r, eval = FALSE}
datelife::make_datelife_query(input = c("Delphinus_delphis",
                                        "Gallus gallus",
                                        "elephas Maximus",
                                        "felis_catus",
                                        "homo-sapiens"))
```
<details>
<summary>make_datelife_query output 1</summary>
```r
#> ... Making a DateLife query.
#> ... Phylo-processing 'input'.
#> 'input' is not a phylogeny.
#> DateLife query done!
#> $cleaned_names
#> [1] "Delphinus delphis" "Gallus gallus"     "elephas Maximus"
#> [4] "felis catus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#> [1] NA
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>

Gives the same result as running:

```{r, eval = FALSE}
datelife::make_datelife_query(input = "Delphinus_delphis,
                                       Gallus gallus,
                                       elephas Maximus,
                                       felis_catus,
                                       homo-sapiens")
```
<details>
<summary>make_datelife_query output 2</summary>
```r
#> ... Making a DateLife query.
#> ... Phylo-processing 'input'.
#> 'input' is not a phylogeny.
#> DateLife query done!
#> $cleaned_names
#> [1] "Delphinus delphis" "Gallus gallus"     "elephas Maximus"
#> [4] "felis catus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#> [1] NA
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>


Note that you can use underscores or spaces to separate binomial scientific names and `datelife` will process them the same.

### b) From a tree with organism taxonomic names as tip labels

You can also provide the taxon names of your organisms of interest as labels on the tips of any phylogenetic tree.
You can provide the tree as a newick string:

```{r, eval = FALSE}
newick <- "((Elephas_maximus,(Homo_sapiens,(Delphinus_delphis, Felis_silvestris))), Gallus_gallus);"
datelife::make_datelife_query(input = newick)
```
<details>
<summary>make_datelife_query output 3</summary>
```r
#> ... Making a DateLife query.
#> ... Phylo-processing 'input'.
#> 'input' is a phylogeny and it is correcly formatted.
#> DateLife query done!
#> $cleaned_names
#> [1] "Elephas maximus"   "Delphinus delphis" " Felis silvestris"
#> [4] " Gallus gallus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Elephas_maximus, Delphinus_delphis, _Felis_silvestris, _Gallus_gallus
#>
#> Rooted; no branch lengths.
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>

You can also provide the tree as a `phylo` object. If you already have a tree as a `phylo` object, you can use that directly. For this example, we will first convert our newick string to `phylo`. You can do this with the function `read.tree` from the `ape` package:

```{r, eval = FALSE}
phylo <- ape::read.tree(text = newick)
datelife::make_datelife_query(input = phylo)
```
<details>
<summary>make_datelife_query output 4</summary>
```r
#> ... Making a DateLife query.
#> ... Phylo-processing 'input'.
#> 'input' is a phylogeny and it is correcly formatted.
#> DateLife query done!
#> $cleaned_names
#> [1] "Elephas maximus"   "Delphinus delphis" "Felis silvestris"
#> [4] "Gallus gallus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Elephas_maximus, Delphinus_delphis, Felis_silvestris, Gallus_gallus
#>
#> Rooted; no branch lengths.
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>

Now save your `datelifeQuery` as an object for the next section:

```{r, eval = FALSE}
my_datelife_query <- datelife::make_datelife_query(input = phylo)
#> ... Making a DateLife query.
#> ... Phylo-processing 'input'.
#> 'input' is a phylogeny and it is correctly formatted.
#> DateLife query done!
```

## 2. Searching the chronogram database

The `datelife_search` function takes a `datelifeQuery` object and goes through the `datelife` chronogram database to find all chronograms matching at least two taxon names:

```{r, eval = FALSE}
datelife::datelife_search(input = my_datelife_query)
```
<details>
<summary>datelife_search default output</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Source chronograms from:
#> 1: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 2: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 3: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 4: Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#> 4 phylogenetic trees
```
</details>

## 3. Summarizing chronogram search results

`datelife_search` will also summarize results from the chronogram database search for you.
By default, it will give you all original chronograms matching the search, as a `phylo` or `multiPhylo` object. You can specify the format you desire for the summary of results with the argument `summary_format`.
There are different summary formats to choose from:

### Citations

`summary_format = "citations"` gives back the study citations from all  original chronograms matching your search:

```{r, eval = FALSE}
citations <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "citations")
```
<details>
<summary>datelife_search output "citations"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#> [1] "Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512"
#> [2] "Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512"
#> [3] "Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512"
#> [4] "Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845"
#> attr(,"datelife_result")
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Delphinus delphis
#> Elephas maximus               0.0             202.6
#> Delphinus delphis           202.6               0.0
#> Felis silvestris            202.6             177.0
#>                   Felis silvestris
#> Elephas maximus              202.6
#> Delphinus delphis            177.0
#> Felis silvestris               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Felis silvestris
#> Elephas maximus               0.0            192.2
#> Felis silvestris            192.2              0.0
#> Delphinus delphis           192.2            173.6
#>                   Delphinus delphis
#> Elephas maximus               192.2
#> Felis silvestris              173.6
#> Delphinus delphis               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Delphinus delphis Felis silvestris
#> Delphinus delphis               0.0            179.6
#> Felis silvestris              179.6              0.0
#> Elephas maximus               221.6            221.6
#>                   Elephas maximus
#> Delphinus delphis           221.6
#> Felis silvestris            221.6
#> Elephas maximus               0.0
#>
#> $`Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845`
#>                   Felis silvestris Delphinus delphis
#> Felis silvestris            0.0000          156.7850
#> Delphinus delphis         156.7850            0.0000
#> Elephas maximus           210.0193          210.0193
#> Gallus gallus             641.0476          641.0476
#>                   Elephas maximus Gallus gallus
#> Felis silvestris         210.0193      641.0476
#> Delphinus delphis        210.0193      641.0476
#> Elephas maximus            0.0000      641.0476
#> Gallus gallus            641.0476        0.0000
#>
#> attr(,"class")
#> [1] "datelifeResult"
#> attr(,"datelife_query")
#> $cleaned_names
#> [1] "Elephas maximus"   "Delphinus delphis" "Felis silvestris"
#> [4] "Gallus gallus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Elephas_maximus, Delphinus_delphis, Felis_silvestris, Gallus_gallus
#>
#> Rooted; no branch lengths.
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>

### MRCA ages

`summary_format = "mrca"` gives back the age of the MRCA of your organisms of interest:

```{r, eval = FALSE}
mrca <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "mrca")
```
<details>
<summary>datelife_search output "mrca"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Source chronograms from:
#> 1: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 2: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 3: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 4: Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#> Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#>                                                                                                                                                                                                                                                       101.3000
#> Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#>                                                                                                                                                                                                                                                        96.1000
#> Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#>                                                                                                                                                                                                                                                       110.8000
#>                                                           Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#>                                                                                                                                                                                                                                                       320.5238
#> attr(,"datelife_result")
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Delphinus delphis
#> Elephas maximus               0.0             202.6
#> Delphinus delphis           202.6               0.0
#> Felis silvestris            202.6             177.0
#>                   Felis silvestris
#> Elephas maximus              202.6
#> Delphinus delphis            177.0
#> Felis silvestris               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Felis silvestris
#> Elephas maximus               0.0            192.2
#> Felis silvestris            192.2              0.0
#> Delphinus delphis           192.2            173.6
#>                   Delphinus delphis
#> Elephas maximus               192.2
#> Felis silvestris              173.6
#> Delphinus delphis               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Delphinus delphis Felis silvestris
#> Delphinus delphis               0.0            179.6
#> Felis silvestris              179.6              0.0
#> Elephas maximus               221.6            221.6
#>                   Elephas maximus
#> Delphinus delphis           221.6
#> Felis silvestris            221.6
#> Elephas maximus               0.0
#>
#> $`Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845`
#>                   Felis silvestris Delphinus delphis
#> Felis silvestris            0.0000          156.7850
#> Delphinus delphis         156.7850            0.0000
#> Elephas maximus           210.0193          210.0193
#> Gallus gallus             641.0476          641.0476
#>                   Elephas maximus Gallus gallus
#> Felis silvestris         210.0193      641.0476
#> Delphinus delphis        210.0193      641.0476
#> Elephas maximus            0.0000      641.0476
#> Gallus gallus            641.0476        0.0000
#>
#> attr(,"class")
#> [1] "datelifeResult"
#> attr(,"datelife_query")
#> $cleaned_names
#> [1] "Elephas maximus"   "Delphinus delphis" "Felis silvestris"
#> [4] "Gallus gallus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Elephas_maximus, Delphinus_delphis, Felis_silvestris, Gallus_gallus
#>
#> Rooted; no branch lengths.
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>

### Newick strings
 
`summary_format = "newick_all"`

```{r, eval = FALSE}
newick_all <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "newick_all")
```
<details>
<summary>datelife_search output "newick_all"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Source chronograms from:
#> 1: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 2: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 3: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 4: Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#> Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#>                                                                                                                                                                                 "((Felis_silvestris:88.5,Delphinus_delphis:88.5):12.8,Elephas_maximus:101.3);"
#> Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#>                                                                                                                                                                                   "((Delphinus_delphis:86.8,Felis_silvestris:86.8):9.3,Elephas_maximus:96.1);"
#> Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#>                                                                                                                                                                                   "((Felis_silvestris:89.8,Delphinus_delphis:89.8):21,Elephas_maximus:110.8);"
#>                                                           Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#>                                                                                                                     "((Elephas_maximus:105.009658,(Felis_silvestris:78.392518,Delphinus_delphis:78.392517):26.617141):215.5141645,Gallus_gallus:320.5238235);"
#> attr(,"datelife_result")
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Delphinus delphis
#> Elephas maximus               0.0             202.6
#> Delphinus delphis           202.6               0.0
#> Felis silvestris            202.6             177.0
#>                   Felis silvestris
#> Elephas maximus              202.6
#> Delphinus delphis            177.0
#> Felis silvestris               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Felis silvestris
#> Elephas maximus               0.0            192.2
#> Felis silvestris            192.2              0.0
#> Delphinus delphis           192.2            173.6
#>                   Delphinus delphis
#> Elephas maximus               192.2
#> Felis silvestris              173.6
#> Delphinus delphis               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Delphinus delphis Felis silvestris
#> Delphinus delphis               0.0            179.6
#> Felis silvestris              179.6              0.0
#> Elephas maximus               221.6            221.6
#>                   Elephas maximus
#> Delphinus delphis           221.6
#> Felis silvestris            221.6
#> Elephas maximus               0.0
#>
#> $`Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845`
#>                   Felis silvestris Delphinus delphis
#> Felis silvestris            0.0000          156.7850
#> Delphinus delphis         156.7850            0.0000
#> Elephas maximus           210.0193          210.0193
#> Gallus gallus             641.0476          641.0476
#>                   Elephas maximus Gallus gallus
#> Felis silvestris         210.0193      641.0476
#> Delphinus delphis        210.0193      641.0476
#> Elephas maximus            0.0000      641.0476
#> Gallus gallus            641.0476        0.0000
#>
#> attr(,"class")
#> [1] "datelifeResult"
#> attr(,"datelife_query")
#> $cleaned_names
#> [1] "Elephas maximus"   "Delphinus delphis" "Felis silvestris"
#> [4] "Gallus gallus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Elephas_maximus, Delphinus_delphis, Felis_silvestris, Gallus_gallus
#>
#> Rooted; no branch lengths.
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>

`summary_format = "newick_sdm"`

```{r, eval = FALSE}
newick_sdm <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "newick_sdm")
```
<details>
<summary>datelife_search output "newick_sdm"</summary>
```r

```
</details>

`summary_format = "newick_median"`

```{r, eval = FALSE}
newick_median <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "newick_median")
```
<details>
<summary>datelife_search output "newick_median"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Trying with overlap = 2
#> Registered S3 method overwritten by 'compare':
#>   method           from
#>   print.comparison testthat
#> Success!
#> ... Calculating a median summary chronogram.
#> Source chronograms from:
#> 1: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 2: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 3: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 4: Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#> [1] "((Elephas_maximus:103.154831,(Delphinus_delphis:87.650002,Felis_silvestris:87.650002)n3:15.504829)n2:217.369003,Gallus_gallus:320.523834)n1:1;"
#> attr(,"datelife_result")
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Delphinus delphis
#> Elephas maximus               0.0             202.6
#> Delphinus delphis           202.6               0.0
#> Felis silvestris            202.6             177.0
#>                   Felis silvestris
#> Elephas maximus              202.6
#> Delphinus delphis            177.0
#> Felis silvestris               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Felis silvestris
#> Elephas maximus               0.0            192.2
#> Felis silvestris            192.2              0.0
#> Delphinus delphis           192.2            173.6
#>                   Delphinus delphis
#> Elephas maximus               192.2
#> Felis silvestris              173.6
#> Delphinus delphis               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Delphinus delphis Felis silvestris
#> Delphinus delphis               0.0            179.6
#> Felis silvestris              179.6              0.0
#> Elephas maximus               221.6            221.6
#>                   Elephas maximus
#> Delphinus delphis           221.6
#> Felis silvestris            221.6
#> Elephas maximus               0.0
#>
#> $`Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845`
#>                   Felis silvestris Delphinus delphis
#> Felis silvestris            0.0000          156.7850
#> Delphinus delphis         156.7850            0.0000
#> Elephas maximus           210.0193          210.0193
#> Gallus gallus             641.0476          641.0476
#>                   Elephas maximus Gallus gallus
#> Felis silvestris         210.0193      641.0476
#> Delphinus delphis        210.0193      641.0476
#> Elephas maximus            0.0000      641.0476
#> Gallus gallus            641.0476        0.0000
#>
#> attr(,"class")
#> [1] "datelifeResult"
#> attr(,"datelife_query")
#> $cleaned_names
#> [1] "Elephas maximus"   "Delphinus delphis" "Felis silvestris"
#> [4] "Gallus gallus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Elephas_maximus, Delphinus_delphis, Felis_silvestris, Gallus_gallus
#>
#> Rooted; no branch lengths.
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>

### `phylo` objects

`summary_format = "phylo_sdm"`

```{r, eval = FALSE}
phylo_sdm <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "phylo_sdm")
```
<details>
<summary>datelife_search output </summary>
```r

```
</details>

`summary_format = "phylo_median"`

```{r, eval = FALSE}
phylo_median <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "phylo_median")
```
<details>
<summary>datelife_search output "phylo_median"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Trying with overlap = 2
#> Success!
#> ... Calculating a median summary chronogram.
#> Source chronograms from:
#> 1: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 2: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 3: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 4: Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Elephas_maximus, Delphinus_delphis, Felis_silvestris, Gallus_gallus
#> Node labels:
#>   n1, n2, n3
#>
#> Rooted; includes branch lengths.
```
</details>

`summary_format = "phylo_all"`

```{r, eval = FALSE}
phylo_all <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "phylo_all")
```
<details>
<summary>datelife_search output "phylo_all"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Source chronograms from:
#> 1: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 2: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 3: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 4: Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#> 4 phylogenetic trees
```
</details>

`summary_format = "phylo_biggest"`

```{r, eval = FALSE}
phylo_biggest <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "phylo_biggest")
```
<details>
<summary>datelife_search output "phylo_biggest"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Source chronograms from:
#> 1: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 2: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 3: Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 4: Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Gallus_gallus, Felis_silvestris, Delphinus_delphis, Elephas_maximus
#>
#> Rooted; includes branch lengths.
```
</details>

### Summary tables

`summary_format = "html"` returns a summary table in "html" format that cns be open and read with any web browser.

```{r, eval = FALSE}
html <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "html")
```
<details>
<summary>datelife_search output "html"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#> [1] "<table border='1'><tr><th>MRCA Age (MY)</th><th>Ntax</th><th>Citation</th><th>Newick</th></tr><tr><td>101.3</td><td>3</td><td>Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512</td><td>((Felis_silvestris:88.5,Delphinus_delphis:88.5):12.8,Elephas_maximus:101.3);</td></tr><tr><td>96.1</td><td>3</td><td>Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512</td><td>((Delphinus_delphis:86.8,Felis_silvestris:86.8):9.3,Elephas_maximus:96.1);</td></tr><tr><td>110.8</td><td>3</td><td>Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512</td><td>((Felis_silvestris:89.8,Delphinus_delphis:89.8):21,Elephas_maximus:110.8);</td></tr><tr><td>320.5238235</td><td>4</td><td>Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845</td><td>((Elephas_maximus:105.009658,(Felis_silvestris:78.392518,Delphinus_delphis:78.392517):26.617141):215.5141645,Gallus_gallus:320.5238235);</td></tr></table>"
#> attr(,"datelife_result")
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Delphinus delphis
#> Elephas maximus               0.0             202.6
#> Delphinus delphis           202.6               0.0
#> Felis silvestris            202.6             177.0
#>                   Felis silvestris
#> Elephas maximus              202.6
#> Delphinus delphis            177.0
#> Felis silvestris               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Elephas maximus Felis silvestris
#> Elephas maximus               0.0            192.2
#> Felis silvestris            192.2              0.0
#> Delphinus delphis           192.2            173.6
#>                   Delphinus delphis
#> Elephas maximus               192.2
#> Felis silvestris              173.6
#> Delphinus delphis               0.0
#>
#> $`Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512`
#>                   Delphinus delphis Felis silvestris
#> Delphinus delphis               0.0            179.6
#> Felis silvestris              179.6              0.0
#> Elephas maximus               221.6            221.6
#>                   Elephas maximus
#> Delphinus delphis           221.6
#> Felis silvestris            221.6
#> Elephas maximus               0.0
#>
#> $`Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845`
#>                   Felis silvestris Delphinus delphis
#> Felis silvestris            0.0000          156.7850
#> Delphinus delphis         156.7850            0.0000
#> Elephas maximus           210.0193          210.0193
#> Gallus gallus             641.0476          641.0476
#>                   Elephas maximus Gallus gallus
#> Felis silvestris         210.0193      641.0476
#> Delphinus delphis        210.0193      641.0476
#> Elephas maximus            0.0000      641.0476
#> Gallus gallus            641.0476        0.0000
#>
#> attr(,"class")
#> [1] "datelifeResult"
#> attr(,"datelife_query")
#> $cleaned_names
#> [1] "Elephas maximus"   "Delphinus delphis" "Felis silvestris"
#> [4] "Gallus gallus"
#>
#> $ott_ids
#> NULL
#>
#> $phy
#>
#> Phylogenetic tree with 4 tips and 3 internal nodes.
#>
#> Tip labels:
#>   Elephas_maximus, Delphinus_delphis, Felis_silvestris, Gallus_gallus
#>
#> Rooted; no branch lengths.
#>
#> attr(,"class")
#> [1] "datelifeQuery"
```
</details>

`summary_format = "data_frame"`

```{r, eval = FALSE}
data_frame <- datelife::datelife_search(input = my_datelife_query,
                          summary_format = "data_frame")
```
<details>
<summary>datelife_search output "data_frame"</summary>
```r
#> ... Running a DateLife search.
#> ... Getting a DateLife result.
#> DateLife result obtained!
#> Input taxa presence across source chronograms:
#>               taxon chronograms
#> 1   Elephas maximus         4/4
#> 2 Delphinus delphis         4/4
#> 3  Felis silvestris         4/4
#> 4     Gallus gallus         1/4
#> Input taxa completely absent from source chronograms:
#>   taxon
#> 1  None
#> DateLife search done!
#>        Age Ntax
#> 1 101.3000    3
#> 2  96.1000    3
#> 3 110.8000    3
#> 4 320.5238    4
#>                                                                                                                                                                                                                                                         Citation
#> 1 Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 2 Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 3 Bininda-Emonds, Olaf R. P., Marcel Cardillo, Kate E. Jones, Ross D. E. MacPhee, Robin M. D. Beck, Richard Grenyer, Samantha A. Price, Rutger A. Vos, John L. Gittleman, Andy Purvis. 2007. The delayed rise of present-day mammals. Nature 446 (7135): 507-512
#> 4                                                           Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845
#>                                                                                                                                     Newick
#> 1                                                             ((Felis_silvestris:88.5,Delphinus_delphis:88.5):12.8,Elephas_maximus:101.3);
#> 2                                                               ((Delphinus_delphis:86.8,Felis_silvestris:86.8):9.3,Elephas_maximus:96.1);
#> 3                                                               ((Felis_silvestris:89.8,Delphinus_delphis:89.8):21,Elephas_maximus:110.8);
#> 4 ((Elephas_maximus:105.009658,(Felis_silvestris:78.392518,Delphinus_delphis:78.392517):26.617141):215.5141645,Gallus_gallus:320.5238235);
```
</details>
