---
title: "package_development"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{package_development}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(datelife)
library(covr)
```

# Coverage of tests

We are using the package `covr`.
We set it up pn Travis, so everytime it tests coverage is calculated and shown very nicely in here.
However, this takes time, because some tests take a looong time.

What I do to check coverage locally:

```{r, eval = FALSE}
cov <- covr::package_coverage()
save(cov, file = "cov.RData")
# This last one will save the object on the root of the package.
```

It is important to add cov.RData to your `.gitignore` file, so the file is not version controlled.

You can load you robejct on any R environment you want (terminal, RStudio, R console)
```{r, eval = TRUE}
load("~/pj_datelife/datelife/cov.RData")
```

Let's look at the object now

```{r, eval = TRUE}
cov
```
If you are using RStudio, the following will print all the lines that are not covered in tests, and will allow you to navigate them interactively `r emo::ji("star_struck")`
```{r, eval = FALSE}
covr::zero_coverage(cov)
```