## Feedback and Information for Developers

We welcome and encourage to post a [GitHub issue](https://github.com/phylotastic/datelife/issues/new)
with any comments, ideas and questions about `datelife`'s software and website.
If you want to contribute with code directly, we welcome and encourage [pull requests](https://github.com/phylotastic/datelife/pulls).

#### _Function documentation_:

Package and function documentation was generated with [roxygen2](https://CRAN.R-project.org/package=roxygen2):

```{r eval = FALSE}
roxygen2::roxygenise()
```

#### _Styling code_:

We used the package [lintr](https://CRAN.R-project.org/package=lintr) to check for coding style:

```{r eval = FALSE}
lintr::lint_package()
```

#### _Calculating test coverage_:

Code coverage was calculated with the package [covr](https://CRAN.R-project.org/package=covr):

```{r eval = FALSE}
cov <- covr::package_coverage()

usethis::use_data(cov, overwrite = TRUE)
```

You can see an interactive report of testing coverage:

```{r eval = FALSE}
covr::report(cov)
```

And, find code with zero coverage:

```{r eval = FALSE}
covr::zero_coverage(cov)
```

#### _Generating `datelife`'s hexsticker_:

Code used to generate current `datelife`'s logo hexsticker is in [data-raw/hexsticker-current.R](https://github.com/phylotastic/datelife/blob/master/data-raw/hexsticker-current.R)

#### _Rendering the vignettes_:

Vignettes are rendered automatically upon built. However, if you wish to see how they look rendered before releasing the package, you can do this with `knitr::knit()`. The following command renders the vignette `Getting_started_with_datelife` as html:


```{r eval = FALSE}
knitr::knit("vignettes/Getting_started_with_datelife.Rmd")
```

#### _Creating a documentation website for the package_

Using `pkgdown` for this is quite straightforward and fun:

```r
usethis::use_pkgdown()
pkgdown::build_site()
```


#### _Releasing to CRAN_

##### _**Updating GitHub actions R CMD check**_

Run the following function from the package [`usethis`](https://usethis.r-lib.org/reference/github_actions.html) to update R CMD Check on GitHub:

```r
usethis::use_github_action_check_standard()
```

This downloads the standard R CMD check workflow from [r-lib action examples](https://github.com/r-lib/actions/blob/v2/examples/check-standard.yaml).

##### _**Local checks**_

To be able to release to CRAN, the first step is to pass the checks locally. To run a local check, you can use the command `R CMD check` from your terminal. For that, change directories to the one above your working clone of the `datelife` repo:

``` bash
cd ../
```

Generate a tar ball for your package by running `R CMD build package-name`:

``` bash
R CMD build datelife
```

Finally, run `R CMD check package-tar-ball` on the tar ball that you just generated:

``` bash
R CMD check --as-cran datelife_0.6.0.tar.gz
```

##### _**Remote checks**_

If you do not have access to different OS to test your package on, the [rhub](https://CRAN.R-project.org/package=rhub) package allows remote testing on a variety of OS with the command:

```r
rhub::check_for_cran()
```

For more `rhub` useful workflows, check out its [documentation](https://r-hub.github.io/rhub/articles/rhub.html). For example

```r
previous_checks <- rhub::list_package_checks(".",
                                  email = "sanchez.reyes.luna@gmail.com",
                                  howmany = 4)
group_check <- rhub::get_check(previous_checks$group[1])
group_check

cran_prep <- check_for_cran()
cran_prep$cran_summary()
```

To submit to CRAN call `devtools::release()` and answer the prompted questions. If the answer to all of these is _yes_, the package will be submitted to CRAN :rocket:

To [check for URL validity](https://blog.r-hub.io/2020/12/01/url-checks/) specifically, use:

```r
devtools::check_win_release()
```
